Binning Strategies and Related Loss for Large Data

### 2015-12-17 versionForJCGS/PaperJCGS.Rnw
Change hex binned tiles to minimally binned rectangular tiles to introduce minimal binning early in simple context to do double duty

\begin{figure}[h]
  \subfloat[Traditional scatterplot ]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/Overplotting.png}}
  \subfloat[Tukey-style open circles]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/OverplottingCircles.png}} 
	\subfloat[Alpha blending]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/OverplottingAlpha.png}}

  \subfloat[Hexagonal binned scatterplot]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/HexBinning.png}}
  \subfloat[Hexagonal binned bubble plot]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/BinBubble.png}}
  \subfloat[Minimally binned scatterplot]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/BinScatter.png}}

@@

 Figure~\ref{fig:scatter-alpha} contains examples of a hexagonally binned scatterplot with frequency encoded as color (d) and a bubble plot with frequency encoded as point size (e). The hexagonally binned scatterplot and the bubble plot are more effective at displaying the shape of the joint density and preserving outliers than any of the scatterplots shown in Figure~\ref{fig:scatter-alpha}(a-c). The bubble plot is prone to suffer from the Hermann-grid illusion \citep{hermann:1870}, where the white spaces between circles on the evenly spaced grid appear shaded due to an optical illusion; whereas this will not occur for a binned scatterplot on a tessellated grid.

A traditional scatterplot is comparable to a \textit{minimally binned scatterplot} -- using small enough bins such that only a single unique coordinate pair exists within each bin --  however, the "bins" of a traditional scatterplot are shaded in a binary manner with no indication of overlapping observations. Alpha blending as used in Figure~\ref{fig:scatter-alpha}(c) extends the binary shading of a standard scatterplot to an implicit shading according to frequency. The shading is implicit because the range of frequency information is not scaled to the range of shading values, so that maximum color saturation is usually reached well before the maximum frequency, truncating the perceivable frequency information. By explicitly shading bins according to frequency, more information is preserved than in a traditional scatter plot, as the frequency domain provides visual weight to tiles which may represent more points.

>>>>

\begin{figure}[hbtp]
  \subfloat[Traditional scatterplot ]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/Overplotting.png}}
  \subfloat[Tukey-style open circles]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/OverplottingCircles.png}} 
	\subfloat[Alpha blending]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/OverplottingAlpha.png}}

  \subfloat[Rectangular minimally binned scatterplot]{\includegraphics[keepaspectratio=TRUE,width=.48\textwidth]{./figure/BinScatter.png}}  
%  \subfloat[Hexagonal binned scatterplot]{\includegraphics[keepaspectratio=TRUE,width=.32\textwidth]{./figure/HexBinning.png}}
  \subfloat[Hexagonal binned bubble plot]{\includegraphics[keepaspectratio=TRUE,width=.48\textwidth]{./figure/BinBubble.png}}

@@

 Figure~\ref{fig:scatter-alpha} contains examples of a rectangular binned scatterplot with frequency encoded as color (d) and a hexagonal binned bubble plot with frequency encoded as point size (e); both of which are more effective at displaying the shape of the joint density and preserving outliers than any of the scatterplots shown in Figure~\ref{fig:scatter-alpha}(a-c). Bubble plots is prone to suffer from the Hermann-grid illusion \citep{hermann:1870}, where the white spaces between circles on the evenly spaced grid appear shaded due to an optical illusion. The rectangular bins in Figure~\ref{fig:scatter-alpha}(d) are one game by one strikeout in size which matches the resolution of the recorded data; this is referred to as a \textit{minimally binned scatterplot}. With only a single unique coordinate pair exists within each bin the tiles are comparable to points in a traditional scatterplot; however, the unique coordinate pairs in a traditional scatterplot are shaded in a binary manner with no indication of overlapping observations. Alpha blending as used in Figure~\ref{fig:scatter-alpha}(c) is akin to bin shading; however the frequency mapping is imperfect because as soon as overlapping points surpass full opacity the perceivable frequency information is truncated. By explicitly shading bins according to frequency, more information is preserved than in a traditional scatter plot, as the frequency domain provides visual weight to tiles which may represent more points.
